---
title: "Analyses of stream E. coli concentrations for: 'Mitigating ecosystem services tradeoffs in semi-arid rangelands by shortening grazing duration in riparian areas'"
author: "Andrew Kleinhesselink"
date: "10/28/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list = ls())

require(tidyverse)
require(scales)
require(gamm4)
require(emmeans)
require(lubridate)
require(cowplot)
require(pbkrtest)
require(stargazer)
require(ggrepel)


```

## Set-up parameters and Labels

Each stream was assigned to one of three treatments: 
- short = "Time-Controlled Rotation" (aka Rapid Rotation)
  - Definition: cattle are moved frequently 
- medium = "Deferred-Rotation" (Four Pasture Rotation)
- long = "Continuous-Turnout" (No rotation)

E. coli thresholds are given by the Utah DWQ and the EPA. 

```{r parameters and labels, include = F}

# Treatment Labels 
# Low = Rapid Rotation/Short Duration/Time-Controlled Rotation
# Medium = Four Pasture Rotation/Medium Duration/Deferred-Rotation
# High = No Rotation/Long Duration/Continuous-Turnout

rotation_labels <- factor( 
  c('Continuous-Turnout', 
  'Deferred-Rotation', 
  'Time-Controlled Rotation'), 
  levels = c('Time-Controlled Rotation', 
             'Deferred-Rotation', 
             'Continuous-Turnout'), ordered = T)

treatment <- factor( c('long', 'medium', 'short') )

treatment_table <- 
  data.frame( Rotation = rotation_labels ,  
              treatment = treatment )
    

# Utah DWQ and EPA thresholds for E. coli. 

threshholds <- data.frame( 
  DOY = Inf, 
  Limit = c(668, 320, 410), 
  Threshold = c('Utah DWQ', 'EPA (75th percentile)', 'EPA (90th percentile)'), 
  `E.coliMPN` = factor( c(668, 320, 410), 
                        levels = c(668, 410, 320), 
                        ordered = T))
# Recreational Window 
# From May 1st to September 30th is the recreational water quality window 
rec_windows <- 
  data.frame( year = c(2016, 2017, 2018)) %>%
  mutate( rec_start_doy = yday(ymd( paste(year, '05', '01', sep = '-') )), 
          rec_end_doy = yday( ymd(paste( year, '09', '30', sep = '-' ))))


xlab <- "Day of year"
ylab <- expression( italic (E) *'.'~ italic( coli )~ '(M.P.N.)')

mycolors <-  hue_pal()(4) 
mycolors2 <- c('darkgray', mycolors[c(4,1)])


```

## Input Data

Read in data from excel spreadsheet "ecoli_data.xlsx". 

```{r data input, include = F, message=F}
stream_data <- 
  readxl::read_xlsx('data/ecoli_data.xlsx') %>% 
  mutate_at(.vars = c('sample_date', 
                      'grazing_start', 
                      'grazing_end', 
                      'trailing_start', 
                      'trailing_end'), .funs = function(x) ymd(x) ) %>%
  mutate( year = year( sample_date))  %>% 
  filter( ! is.na(grazing_start)) # Remove pasture with no grazing. 

stream_data <- 
  stream_data %>% 
  group_by( year, stream, pasture, treatment ) %>% 
  mutate( Cattle = factor(cattle_present, levels = c(0, 1), labels = c('Absent', 'Present'))) %>%
  mutate( DOY = yday(ymd(sample_date))) %>% 
  mutate( log_Y = log10( ecoli_MPN)) %>% 
  ungroup() %>% 
  mutate( treatment = factor(treatment)) %>% 
  left_join( treatment_table ) %>% 
  mutate( cattle_lab = ifelse( Cattle == 'Present', '*', '' )) %>% 
  distinct() 

cattle_windows <- 
  stream_data %>% 
  distinct(year, stream, pasture, treatment, Rotation, grazing_start, grazing_end, trailing_start, trailing_end ) %>% 
  mutate( grazing_start = yday( grazing_start), grazing_end = yday( grazing_end), trailing_start = yday(trailing_start), trailing_end = yday(trailing_end)) %>% 
  mutate( y_low = 0, y_hi = Inf) %>% 
  mutate( DOY = 100, ecoli_MPN = 1 )

```

## Construct Example System

This step sets up a reference grazing calandar for predictions. Each treatment is given a start DOY and a duration and the end DOY is then calculated. This is mostly for plotting but can also be used to project number of days of E. coli levels above thresholds for an example system. 

```{r Example System}

# Set grazing start (DOY)
start_doy <- 150

# Set grazing durations (ndays):
short_duration <- 20
medium_duration <- 50
long_duration <- 120

ref_grazing_calandar <- 
  stream_data %>% 
  ungroup() %>% 
  distinct(treatment) %>% 
  mutate( grazing_start = start_doy) %>% 
  mutate( grazed_period = NA) %>% 
  mutate( grazed_period = ifelse(treatment == 'medium', medium_duration, grazed_period)) %>% 
  mutate( grazed_period = ifelse(treatment == 'short' , short_duration, grazed_period)) %>% 
  mutate( grazed_period = ifelse(treatment == 'long', long_duration, grazed_period)) %>% 
  mutate( grazing_end = as.numeric( grazing_start + grazed_period)) 


pred_grid <- expand.grid( treatment = levels( stream_data$treatment), 
                          DOY = min(stream_data$DOY):max(stream_data$DOY)) %>% 
  left_join( stream_data %>% 
               ungroup %>% 
               distinct( treatment, Rotation) ) 

pred_grid <- 
  pred_grid %>% 
  group_by( treatment, Rotation ) %>% 
  arrange( treatment, Rotation, DOY) %>% 
  left_join(
    ref_grazing_calandar  )  %>% 
  mutate( period = 'pre') %>% 
  mutate( period = ifelse( DOY > grazing_start , 'grazed', period )) %>% 
  mutate( period = ifelse( DOY > grazing_end, 'post', period )) %>% 
  mutate( Cattle = factor(ifelse( period == 'grazed', 'Present', 'Absent' ))) 

pred_grid_windows <- 
  pred_grid %>% 
  group_by(treatment, Rotation, Cattle ) %>% 
  filter( Cattle == 'Present' ) %>% 
  summarise( DOY1 = min(DOY), DOY2 = max(DOY)) %>% 
  mutate( DOY = 200, ecoli_MPN = 100)

```

##  Calendar plots 

Plots show daily E. coli measurements in each stream each year.  Plots are faceted by pasture and stream and labeled with treatment.  Different years given different colors. Asterisks show days when cattle were present. 

```{r, calendar plots }
stream_data$stream_label <- paste0( "Stream=", stream_data$stream)
stream_data$pasture_label <- paste0( "Pasture=", str_to_title(stream_data$pasture))

threshholds$Limit

fig_ts <- 
  stream_data %>% 
  ggplot( aes( x = DOY, y = ecoli_MPN , color = factor(year) )) + 
  geom_vline(aes(xintercept = 121), color = c('darkgray'), linetype = c(1)) + 
  geom_vline(aes(xintercept = 273), color = c('darkgray'), linetype = c(1)) + 
  geom_hline(data = threshholds, 
             aes( x = DOY, yintercept = Limit, linetype = E.coliMPN), show.legend = T) +  
  geom_point() + 
  geom_line() + 
  geom_text( aes( label = cattle_lab ), nudge_y = 0.1, show.legend = F) + 
  scale_y_log10() + 
  scale_color_manual(values = mycolors2) + 
  facet_wrap( Rotation ~ pasture_label + stream_label) 

# show figure 

fig_ts +  
    theme_classic() + 
    xlim( c(100, 300)) + 
    scale_linetype(name = 'Regulatory limits') + 
    scale_color_discrete(name = "Year") + 
    ylab(ylab) + 
    xlab(xlab)

```


```{r Save Figure 1}

ggsave("Figure_1_annual_timeseries.png", 
       fig_ts +  
        theme_classic() + 
        xlim( c(100, 300)) + 
        scale_linetype(name = 'Regulatory limits') + 
        scale_color_discrete(name = "Year") + 
        ylab(ylab) + 
        xlab(xlab) , width = 9, height = 6, units = 'in'
)

```

## Model Fitting 
Fit a series of generalized additive models with hierarchical error structure. These are fitted to the E. coli concentrations (Log-base-10).

Independent variables: 
- Treatment/Rotation (short, medium, long)
- Cattle (present/absent) 
- DOY (1-365) 

Grouping factors for error: 
- stream 
- year 
- pasture

```{r gamm2}
gam1 <- gamm4(data = stream_data , log_Y ~ treatment + Cattle + s(DOY, bs = 'cr', by = treatment), random = ~ (1|year:stream:pasture) + (1|stream))

gam2 <- gamm4(data = stream_data , log_Y ~ treatment + Cattle + s(DOY, bs = 'cr'), random = ~ (1|year:stream:pasture) + (1|stream))

gam3 <- gamm4(data = stream_data , log_Y ~ treatment + s(DOY, bs = 'cr', by = treatment), random = ~ (1|year:stream:pasture) + (1|stream))

gam4 <- gamm4(data = stream_data , log_Y ~ Cattle + s(DOY, bs = 'cr'), random = ~ (1|year:stream:pasture) + (1|stream))

gam5 <- gamm4(data = stream_data , log_Y ~ treatment + s(DOY, bs = 'cr'), random = ~ (1|year:stream:pasture) + (1|stream))

gam_null <- gamm4( data = stream_data, log_Y ~ s(DOY, bs = 'cr'), random = ~( 1|year:stream:pasture) + (1|stream))

# Model comparison: 

anova(gam1$mer, gam2$mer)

anova(gam2$mer, gam3$mer)

anova(gam2$mer, gam4$mer)

anova(gam1$mer, gam2$mer, gam3$mer, gam4$mer, gam5$mer, gam_null$mer)

```

## Model Comparison Tables 

```{r}

mod_list <- list( gam_null, gam1, gam2, gam3, gam4, gam5)

my_forms <- as.character( lapply( mod_list, function(x) formula(  x$gam )  ) )

mod_summary_table <- 
  data.frame(  anova(gam1$mer, gam2$mer, gam3$mer, gam4$mer, gam5$mer, gam_null$mer) ) %>% 
  mutate( Model = row.names(.) ) %>% 
  arrange( Model) %>% 
  mutate( `Fixed effects` = my_forms) %>%
  rename( "DF" = Df) %>% 
  dplyr::select( Model, `Fixed effects`, DF, logLik, AIC) %>% 
  arrange( AIC ) %>% 
  mutate( `Random effects` = "~(1|year:stream:pasture) + (1|stream)") %>% 
  mutate( `Random effects` = ifelse( row_number() != 1, '""', `Random effects` ))  %>% 
  mutate( `Fixed effects` = str_trim(str_remove_all(`Fixed effects`, '(log_Y )|(, bs = "cr")'))) %>% 
  mutate( Model = str_remove_all(Model, '[(gam)(mer)_]|(\\$)' )) %>% 
  mutate( logLik = round( logLik, 1), AIC = round(AIC, 1))


stargazer(mod_summary_table, 
          summary = F, 
          type = 'html', 
          out = 'table_1_model_description.html', 
          digits = 1, rownames = F) 

```

## Model Selection 

Select model with lowest AIC and fewest parameters. 

```{r}
top_gam <- gam2  ### best model based on combination of AIC and DF 
summary( top_gam$mer)  

anova(top_gam$gam, freq = F)

test_cattle <- KRmodcomp(top_gam$mer, gam5$mer)
test_treatment <- KRmodcomp( top_gam$mer, gam4$mer)

print( 
  summary(test_treatment), 
  summary(test_cattle)
)

```

## GAMM predictions on "example" season

```{r}
yhat <- predict(top_gam$gam, newdata = pred_grid, type = 'response', se.fit = T)
pred_grid$yhat.top <- 10^( yhat$fit ) 
pred_grid$lcl.top <- 10^( yhat$fit - 2*yhat$se.fit) 
pred_grid$ucl.top  <- 10^( yhat$fit + 2*yhat$se.fit)

yhat_pred <- predict(top_gam$gam, newdata = pred_grid, type = 'response', se.fit = T, interval = 'predict')
pred_grid$y_predict.top <- 10^( yhat_pred$fit ) 
pred_grid$lcl.top <- 10^( yhat$fit - 2*yhat$se.fit) 
pred_grid$ucl.top  <- 10^( yhat$fit + 2*yhat$se.fit)

yhat <- predict(gam1$gam, newdata = pred_grid, type = 'response', se.fit = T)
pred_grid$yhat.gam1 <- 10^( yhat$fit ) 
pred_grid$lcl.gam1 <- 10^( yhat$fit - 2*yhat$se.fit) 
pred_grid$ucl.gam1  <- 10^( yhat$fit + 2*yhat$se.fit)

yhat <- predict(gam3$gam, newdata = pred_grid, type = 'response', se.fit = T)
pred_grid$yhat.gam3 <- 10^( yhat$fit ) 
pred_grid$lcl.gam3 <- 10^( yhat$fit - 2*yhat$se.fit) 
pred_grid$ucl.gam3  <- 10^( yhat$fit + 2*yhat$se.fit)

yhat <- predict(gam4$gam, newdata = pred_grid, type = 'response', se.fit = T)
pred_grid$yhat.gam4 <- 10^( yhat$fit ) 
pred_grid$lcl.gam4 <- 10^( yhat$fit - 2*yhat$se.fit) 
pred_grid$ucl.gam4 <- 10^( yhat$fit + 2*yhat$se.fit)

yhat <- predict(gam5$gam, newdata = pred_grid, type = 'response', se.fit = T)
pred_grid$yhat.gam5 <- 10^( yhat$fit ) 
pred_grid$lcl.gam5 <- 10^( yhat$fit - 2*yhat$se.fit) 
pred_grid$ucl.gam5 <- 10^( yhat$fit + 2*yhat$se.fit)

yhat <- predict(gam_null$gam, newdata = pred_grid, type = 'response', se.fit = T)
pred_grid$yhat.gam.null <- 10^( yhat$fit ) 
pred_grid$lcl.gam.null <- 10^( yhat$fit - 2*yhat$se.fit) 
pred_grid$ucl.gam.null <- 10^( yhat$fit + 2*yhat$se.fit)

```

## Plot data and model fit

Construct baseline figure and add data and prediction to it. 
```{r baseline figure}

fig2 <- 
  stream_data %>% 
  distinct(Rotation, treatment, DOY, year, stream, pasture, ecoli_MPN ) %>% 
  ggplot( aes( y = ecoli_MPN, x = DOY, color = Rotation, fill = Rotation)) + 
  geom_vline(aes(xintercept = 121), color = c('darkgray'), linetype = c(1)) + 
  geom_vline(aes(xintercept = 273), color = c('darkgray'), linetype = c(1)) + 
  geom_hline(data = threshholds, 
             aes( yintercept = Limit, linetype = `E.coliMPN` )) +
  scale_color_manual(values = mycolors2, 
                     guide = guide_legend( title.position = 'top', 
                                           title.hjust = 0.5)) +
  scale_linetype_manual( values = c(1:3)) + 
  scale_y_log10() + 
  theme_classic() + 
  theme(legend.position = 'top')


``` 
## Plot model fits 

```{r }

fig2gam.top <- 
  fig2 + 
  geom_point( color = NA) + 
  geom_ribbon(data = pred_grid, aes( x = DOY,  y = yhat.top, ymin = lcl.top, ymax = ucl.top, fill = Rotation, color = NULL), alpha = 0.4 ) + 
  geom_line( data = pred_grid, aes( x = DOY, y = yhat.top, color = Rotation), alpha = 1, size = 1.5) + 
  geom_rect( data = pred_grid_windows, aes(xmin = DOY1, xmax = DOY2, ymin = 0, ymax = Inf), alpha = 0.2, color = NA, fill = 'gray') + 
  scale_fill_manual(values = mycolors2, guide = guide_legend(title.position = 'top')) + 
  scale_linetype_manual( guide = "none" , values = c(1:3)) + 
  facet_wrap( ~ Rotation , ncol = 1) + 
  xlim( 100, 300 )

fig2gam1 <- 
  fig2 + 
  geom_point( color = NA) + 
  geom_ribbon(data = pred_grid, aes( x = DOY,  y = yhat.gam1, ymin = lcl.gam1, ymax = ucl.gam1, fill = Rotation, color = NULL), alpha = 0.4 ) + 
  geom_line( data = pred_grid, aes( x = DOY, y = yhat.gam1, color = Rotation), alpha = 1, size = 1.5) + 
  geom_rect( data = pred_grid_windows, aes(xmin = DOY1, xmax = DOY2, ymin = 0, ymax = Inf), alpha = 0.2, color = NA, fill = 'gray') + 
  scale_fill_manual(values = mycolors2, guide = guide_legend(title.position = 'top')) + 
  scale_linetype_manual( guide = "none" , values = c(1:3)) + 
  theme(axis.title.y = element_blank()) + 
  facet_wrap( ~ Rotation , ncol = 1) + 
  xlim( 100, 300 )

fig2gam3 <- 
  fig2 + 
  geom_point( color = NA) + 
  geom_ribbon(data = pred_grid, aes( x = DOY,  y = yhat.gam3, ymin = lcl.gam3, ymax = ucl.gam3, fill = Rotation, color = NULL), alpha = 0.4 ) + 
  geom_line( data = pred_grid, aes( x = DOY, y = yhat.gam3, color = Rotation), alpha = 1, size = 1.5) + 
  geom_rect( data = pred_grid_windows, aes(xmin = DOY1, xmax = DOY2, ymin = 0, ymax = Inf), alpha = 0.2, color = NA, fill = 'gray') + 
  scale_fill_manual(values = mycolors2, guide = guide_legend(title.position = 'top')) + 
  scale_linetype_manual( guide = "none" , values = c(1:3)) + 
  theme(axis.title.y = element_blank()) + 
  facet_wrap( ~ Rotation , ncol = 1) + 
  xlim( 100, 300 )

fig2gam4 <- 
  fig2 + 
  geom_point( color = NA) + 
  geom_ribbon(data = pred_grid, aes( x = DOY,  y = yhat.gam4, ymin = lcl.gam4, ymax = ucl.gam4, fill = Rotation, color = NULL), alpha = 0.4 ) + 
  geom_line( data = pred_grid, aes( x = DOY, y = yhat.gam4, color = Rotation), alpha = 1, size = 1.5) + 
  geom_rect( data = pred_grid_windows, aes(xmin = DOY1, xmax = DOY2, ymin = 0, ymax = Inf), alpha = 0.2, color = NA, fill = 'gray') + 
  scale_fill_manual(values = mycolors2, guide = guide_legend(title.position = 'top')) + 
  scale_linetype_manual( guide = "none" , values = c(1:3)) + 
  theme(axis.title.y = element_blank()) + 
  facet_wrap( ~ Rotation , ncol = 1) + 
  xlim( 100, 300 )

plot_grid(fig2gam.top + ggtitle('top gam'),  
          fig2gam1 + ggtitle('gam1'), 
          fig2gam4 + ggtitle('gam4'),  
          fig2gam3 + ggtitle('gam3'), nrow = 1)

fig2gam.top + ylab(ylab) + xlab( xlab)

```

## Save plot of top model 

```{r}

ggsave( fig2gam.top + 
          ylab(ylab) + 
          xlab( xlab), 
        filename = 'figure_2_representative_grazing_season.png', 
        width = 6, height = 6, units = 'in')

```

## Plot predicted vs. observed 

```{r}
par(mfrow = c( 1, 3 ))
plot( predict(top_gam$gam), stream_data$log_Y)
abline(a = 0, b = 1)

plot( predict(gam1$gam), stream_data$log_Y)
abline(a = 0,b = 1)

plot( predict( gam3$gam), stream_data$log_Y)
abline(a = 0,b = 1)
par( mfrow = c(1,1))

```

## GAMM summary 
```{r}

summary( top_gam$gam) 

```

## Summary of hierarchical error 
```{r}

summary(top_gam$mer)

```

## Analyze days exceeding thresholds 

Find the number of daily **E. coli** measures above regulatory thresholds for each stream in each treatment level.  For example season, use average **E. coli** concentrations predicted by the model.  Only count days within the recreational water quality window. 

```{r calculate days exceeding thresholds}

days_exceeding <- 
  pred_grid %>% 
  filter( DOY >= rec_windows$rec_start_doy[ rec_windows$year == 2017], 
          DOY <= rec_windows$rec_end_doy[ rec_windows$year == 2017]) %>%
  distinct( treatment, DOY, yhat.top) %>% 
  group_by( treatment) %>% 
  summarise( `320` = sum(yhat.top > 320), 
             `410` = sum(yhat.top > 410), 
             `668`= sum(yhat.top > 668)) %>% 
  left_join(stream_data %>% 
              ungroup() %>% 
              distinct(treatment, Rotation )) %>% 
  select( Rotation, `320`:`668`) %>%  
  gather( Limit, ndays, `320`:`668`) %>% 
  spread( Rotation, ndays) %>% 
  mutate( Limit = as.numeric(Limit)) %>% 
  left_join(threshholds)

stargazer(days_exceeding %>% 
            select(Threshold, Limit,
                   `Time-Controlled Rotation`:`Continuous-Turnout`),
          summary = F, 
          type = 'html', 
          out = 'table_2_predicted_days_exceeding.html', 
          digits = 1, rownames = F)

stream_data %>% 
  group_by( Rotation ) %>% 
  summarise( `ndays above 320 MPN` = sum(ecoli_MPN > 320)/n(), 
             `ndays above 410 MPN` = sum(ecoli_MPN > 410)/n(), 
             `ndays above 668 MPN`= sum(ecoli_MPN > 668)/n())


```

## Number of days daily E. coli predicted to be above each threshhold

Calculate the number of days per recreation season that **E. coli** concentrations are predicted to be above Utah DWQ thresholds or EPA 75th and 95th percentiles.  Calculate median and 95% Credible Intervals by re-sampling from 10000 daily predictions from posterior predictive distribution. 

```{r } 
set.seed(1) # for Monte Carlo reproduciblility

pred_ndays_above <- function(gam_mod, new_data, threshholds){ 
  # based on advice from Simon Wood https://stat.ethz.ch/pipermail/r-help/2011-April/275632.html
  beta <- coef(gam_mod$gam)
  Vb <- vcov(gam_mod$gam)

  ## Simulate replicate beta vectors from posterior...
  Cv <- chol(Vb)
  n.rep=10000
  nb <- length(beta)
  br <- t(Cv) %*% matrix(rnorm(n.rep*nb),nb,n.rep) + beta

  ## Generate replicate linear predictors...
  Xp <- predict(gam_mod$gam, newdata = new_data, type="lpmatrix")
  lp <- Xp%*%br

  gam_summary <- ( summary( gam_mod$gam ) )
  est_scale <- sqrt( gam_summary$scale)

  ## simulate from normal deviates with mean and scale
  yr <- matrix(
    rnorm(length(lp),lp,sd = est_scale),
    nrow(lp),
    ncol(lp)
    )

  ## compute 95% prediction interval...
  PI <- apply(yr,1,quantile,prob=c(.025,0.975))

  new_data$lpi <- 10^PI['2.5%', ]
  new_data$upi <- 10^PI['97.5%', ]

  pred_split <- new_data %>% 
    select( treatment, DOY, Rotation, Cattle) %>% 
    bind_cols(sim = yr) %>% 
    split( f = .$treatment ) 

  days_above_table <- 
    expand.grid( 
      treatment = c('long', 'medium', 'short'), 
      Limit = threshholds$Limit, 
      min = NA, 
      q2.5 = NA, 
      q50 = NA, 
      q97.5 = NA, 
      max = NA, 
      n = NA)

  for( i in 1:nrow(days_above_table)){ 
    
    temp_sims <- 
      pred_split[[days_above_table$treatment[i]]]$sim
    
    days_above_table[i, c('min', 'q2.5', 'q50', 'q97.5', 'max')] <- 
      quantile( colSums( (10^temp_sims)  > days_above_table$Limit[i] ), 
                c( 0, 0.025, 0.5, 0.975, 1)) %>% round()
    
    days_above_table[i, 'n' ] <- 
      nrow(temp_sims)
      
  }
  
  return(days_above_table)
}

days_above_pred <- 
  pred_ndays_above(top_gam, 
                   stream_data %>% 
                     left_join(rec_windows) %>%
                     filter( DOY >= rec_start_doy, 
                             DOY <= rec_end_doy), 
                   threshholds) %>% 
  gather( stat, val, min:max) %>%
  rename( 'days_above' = val) %>% 
  mutate( pdays_above = days_above/n) %>% 
  mutate( type = 'predicted')

```

```{r }
observed_days <- 
  stream_data %>% 
  left_join(rec_windows) %>%
  filter( DOY >= rec_start_doy,  
          DOY <= rec_end_doy) %>% 
  select( Rotation, ecoli_MPN) %>%
  group_by(Rotation ) %>% 
  summarise( mpn_320 = sum(ecoli_MPN > 320), 
             mpn_410 = sum(ecoli_MPN > 410), 
             mpn_668 = sum(ecoli_MPN > 668), n = n())  %>% 
  gather( thresh, days_above, starts_with('mpn')) %>%
  rename( 'Limit' = thresh) %>% 
  mutate( Limit = as.numeric(str_extract(Limit, '\\d+' ))) %>% 
  mutate( pdays_above = days_above/n) %>% 
  mutate( type = 'observed') %>% 
  mutate( stat = 'q50')

days_above_comparison <- 
  days_above_pred %>% 
    mutate(Rotation = 
           factor(treatment,
                  labels = 
                    c('Continuous-Turnout',
                      'Deferred-Rotation',
                      'Time-Controlled Rotation'))) %>% 
  bind_rows(observed_days %>% 
              mutate( Rotation = factor( Rotation, ordered = F))) %>% 
    mutate( 
      days_above_label = paste0( days_above, '/', n))

```

  
```{r }
cred_intervals <- 
  days_above_comparison %>% 
  filter( type == 'predicted', Limit == '668' ) %>% 
  filter( stat %in% c('q2.5', 'q97.5')) %>% 
  select( Rotation, stat, pdays_above, type ) %>% 
  spread( stat, pdays_above)

days_exceeding_plot <- 
  days_above_comparison %>%
  filter( Limit == 668, stat == 'q50') %>%
  select( -treatment) %>% 
  left_join(cred_intervals) %>%
  ggplot( aes( x = Rotation, 
               y = pdays_above, 
               color = type, 
               shape = type )) + 
  geom_point(position = position_dodge( width = 0.5 ), size = 2) + 
  geom_errorbar(aes( x = Rotation,
                     ymin = q2.5, ymax = q97.5, color = type, shape = type), 
                position = position_dodge(width = 0.5), size = 0.5, width = 0.5) +
  geom_text_repel(
    aes(label = days_above_label),
    hjust = 0,
    position = position_dodge(width = 0.5), 
    show.legend = F) +
  scale_shape_manual(values = c(2,1)) + 
  theme_bw() + 
  theme(legend.title = element_blank(), 
        axis.title.x = element_blank(), 
        panel.grid = element_blank()) + 
  ylab( 'Proportion of Days Above\nUtah DWQ Limit' ) 
  

ggsave(days_exceeding_plot, 
       filename = 'figure_4_days_above_DWQ.png', 
       height = 5, 
       width = 6, 
       units = 'in')

```

### Make a table comparing days above thresholds 

```{r }
days_above_comparison %>% 
  filter( type == 'predicted') %>% 
  select( Limit:Rotation, -n, -days_above, -type) %>% 
  spread( stat, pdays_above)  %>% 
  left_join(
    days_above_comparison %>% 
      filter( type == 'observed') %>% 
      select( Limit:Rotation, -n, -days_above, -type) %>% 
      spread( stat, pdays_above) %>% 
      rename( 'observed' = q50)
  ) %>% 
  write_csv(path = 'prop_days_above_comparison.csv')

```

## Plot top model 
```{r}
saveRDS( top_gam, file = 'Grazing-Windows/top_gam.rds')

save(fig2, pred_grid, mycolors2, ylab, xlab, pred_ndays_above, threshholds, rec_windows, file = 'Grazing-Windows/model_data.rda') # save model for shiny App  

```


## Marginal effect of cattle and treatment on E. coli, removing DOY effect

```{r fig1, echo=FALSE}

pred_grid_cattle_present <- 
  pred_grid %>% 
  filter( DOY == 211 ) %>% 
  mutate( Cattle = 'Present') %>%
  select( treatment, DOY, Rotation, Cattle)  

pred_grid_cattle_absent <- 
  pred_grid %>% 
  filter( DOY == 211 ) %>%
  mutate( Cattle = 'Absent') %>% 
  select( treatment, DOY, Rotation, Cattle )

pred_grid_CATTLE <- bind_rows(pred_grid_cattle_present, pred_grid_cattle_absent)
Xp <- predict( top_gam$gam, newdata = pred_grid_CATTLE, type = "lpmatrix" )[, 1:4] # remove smoothing terms 

a  <- rep(1, 6)
Xs <- t(a) %*% Xp 
Vb <- as.matrix( vcov(top_gam$gam)[1:4, 1:4])
v <- NA
for( i in 1:nrow( Xp)){ 
  v[i] <- Xp[i,] %*% Vb %*% Xp[i,]
}
se <- sqrt(v)
my_coefs <- coef(top_gam$gam)

mean_cattle_absent <- mean( log10( stream_data$ecoli_MPN[ stream_data$Cattle == 'Absent' ]  ) )

my_coefs[1] <- - my_coefs[3] # Reset intercept so short treatment is reference level   

mu <- Xp %*% my_coefs[1:4] # exclude smoothing terms 

pred_grid_CATTLE$yhat <- mu[,1]
pred_grid_CATTLE$ul <- mu[,1] + sqrt( se )
pred_grid_CATTLE$ll <- mu[,1] - sqrt( se )

# Plot marginal effects of treatment and Cattle 

ylab_marginal <- expression( Marginal~effect~'('*Log[10]~italic(E)*'.'~italic(coli)*' '%+-%'S.E.'*')')

treatment_effects <- 
  pred_grid_CATTLE %>% 
  ggplot( aes( x = Rotation, y = yhat, color = Cattle) ) + 
  geom_point(position = position_dodge(width = 1)) + 
  geom_errorbar(aes( ymin = ll, ymax = ul), position = position_dodge(width = 1)) + 
  ylab(ylab_marginal) + 
  scale_color_manual(values = c('#00BFC4', '#F8766D') ) + 
  theme_classic() +
  theme(axis.title.x = element_blank() ) + 
  geom_hline( aes( yintercept = 0 ), linetype = 2, color = "gray") 


ggsave(treatment_effects, 
       filename = 'figure_3_treatment_effects.png', 
       width = 6, 
       height = 3, 
       units = 'in')

```

## Original GAMM 

For reference, we recreate the original GAMM analysis conducted by Jereme Gaeta for Kris Hulvey.  This is being replaced by the new analysis and figures above. 

```{r gamm }

gam_original <- 
  gamm4(data = stream_data , 
        log_Y ~ treatment + s(DOY, bs = 'cr', by = treatment), 
        random = ~ (1|stream:pasture))

yhat <- 
  predict(gam_original$gam, 
          newdata = pred_grid, 
          type = 'response', 
          se.fit = T)

pred_grid$yhat.gam_original <- 10^( yhat$fit ) 
pred_grid$lcl.gam_original <- 10^( yhat$fit - 2*yhat$se.fit) 
pred_grid$ucl.gam_original <- 10^( yhat$fit + 2*yhat$se.fit)


```

## Original Figure 2(a)
```{r fig2a}

fig2a <- 
  fig2 + 
  geom_point(alpha = 0.7, size = 4) + 
  geom_point(alpha = 1, size = 4, shape = 1)



```

## Original Figure 2(b)

```{r}

fig2b <-
  fig2 +
  geom_point(color = NA) +
  geom_ribbon(
  data = pred_grid,
  aes(
  x = DOY,
  y = yhat.gam_original,
  ymin = lcl.gam_original,
  ymax = ucl.gam_original,
  fill = Rotation,
  color = NULL
  ),
  alpha = 0.4
  ) +
  geom_line(
  data = pred_grid,
  aes(x = DOY, y = yhat.gam_original, color = Rotation),
  alpha = 1,
  size = 1.5
  ) +
  scale_fill_manual(values = mycolors2, guide = guide_legend(title.position = 'top')) +
  scale_linetype_manual(guide = "none" , values = c(1:3)) +
  theme(axis.title.y = element_blank())

```

## Original Figure 2 combined 
```{r}

prow <- plot_grid(
  fig2a + theme(legend.position="none"),
  fig2b + theme(legend.position="none"),
  align = 'vh',
  labels = c("A", "B"),
  hjust = -1,
  nrow = 1
)

legend <- get_legend(
  # create some space to the left of the legend
  fig2b + theme(legend.box.margin = margin(0, 0, 0, 12))
)


plot_grid(legend, prow,  ncol = 1, rel_heights = c(0.1, 1))

```

